---
title: "HW4"
author: "Ben Hertzberg"
date: "2022-11-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
titanic <- read.csv('./data/titanic.csv')
library(tidymodels)
library(tidyverse)
library(ggplot2)
library(discrim)
library(poissonreg)
library(corrr)
library(klaR)
tidymodels_prefer()
```


```{r}
titanic$survived <- as.factor(titanic$survived)
titanic$pclass <- as.factor(titanic$pclass)
```


### Question 1

```{r}
set.seed(619)

ttnc_split <- initial_split(titanic, prop = 0.80, strata = survived)
ttnc_train <- training(ttnc_split)
ttnc_test <- testing(ttnc_split)

```

```{r}
dim(ttnc_train)
dim(ttnc_test)
```
712 training observations, 179 testing observations, which is about 80% training and 20% testing.

```{r}
ttnc_rec <- recipe(survived ~ pclass + sex + age + sib_sp + parch + fare, data = ttnc_train) %>%

  step_impute_linear() %>%
  step_dummy(all_nominal_predictors()) %>%  
  step_normalize(all_numeric_predictors()) %>%
  step_interact(terms = ~starts_with("sex"):fare) %>%
  step_interact(terms = ~age:fare)
```

### Question 2

```{r}
ttnc_folds <- vfold_cv(ttnc_train, v = 10)
ttnc_folds
```


### Question 3

In your own words, explain what we are doing in Question 2. What is *k*-fold cross-validation? Why should we use it, rather than simply fitting and testing models on the entire training set? If we **did** use the entire training set, what resampling method would that be?

In Q2, we randomly assigned observations from the training data into 10 roughly equal subsets to be used in k-fold cross-validation. K-fold CV offers a way to check how well our model is doing before applying it on the testing data. This is done by training a model on every fold except one, and then effectively testing on that last fold. K total models are trained, each one using a different fold to test on and training on all the others. This offers the benefit of checking how a model works on data it was not trained on before actually applying it to the test data. If we fit and train models on the entire training set right away, we have to use the testing data to do this check, and then we can't make adjustments to the model without risking data leakage.

### Question 4

```{r}
log_reg <- logistic_reg() %>% 
  set_engine('glm') %>% 
  set_mode('classification')

log_wkflow <- workflow() %>% 
  add_model(log_reg) %>% 
  add_recipe(ttnc_rec)

lda_mod <- discrim_linear() %>% 
  set_engine('MASS') %>% 
  set_mode('classification')

lda_wkflow <- workflow() %>% 
  add_model(lda_mod) %>% 
  add_recipe(ttnc_rec)

qda_mod <-discrim_quad() %>% 
  set_engine('MASS') %>% 
  set_mode('classification')

qda_wkflow <- workflow() %>% 
  add_model(qda_mod) %>% 
  add_recipe(ttnc_rec)
```

I am fitting 3 models to 10 folds, which leads to a total of 30 models.

### Question 5

```{r, eval = FALSE}
log_fit_k <- fit_resamples(log_wkflow, ttnc_folds)
lda_fit_k <- fit_resamples(lda_wkflow, ttnc_folds)
qda_fit_k <- fit_resamples(qda_wkflow, ttnc_folds)
```

```{r, eval = FALSE}
write_rds(log_fit_k, file = 'SavedModels/log_res.rds')
write_rds(lda_fit_k, file = 'SavedModels/lda_res.rds')
write_rds(qda_fit_k, file = 'SavedModels/qda_res.rds')
```

```{r}
log_res <- read_rds(file = 'SavedModels/log_res.rds')
lda_res <- read_rds(file = 'SavedModels/lda_res.rds')
qda_res <- read_rds(file = 'SavedModels/qda_res.rds')
```


### Question 6

```{r}
collect_metrics(log_res)
collect_metrics(lda_res)
collect_metrics(qda_res)
```

QDA performed the worst. LDA and Logistic Regression performed very similarly, but LDA just barely has a higher mean accuracy and lower standard error. Therefore, the LDA model performed the best.

### Question 7

Now that you’ve chosen a model, fit your chosen model to the entire training dataset (not to the folds).

### Question 8

Finally, with your fitted model, use `predict()`, `bind_cols()`, and `accuracy()` to assess your model’s performance on the testing data!

Compare your model’s testing accuracy to its average accuracy across folds. Describe what you see.
