---
title: "HW5"
author: "Ben Hertzberg"
date: "2022-11-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(janitor)
library(ggplot2)
library(tidymodels)
library(tidyverse)
library(glmnet)
```


### Exercise 1

```{r}
pok <- read.csv('./data/Pokemon.csv')
library(janitor)
```

```{r}
pok <- clean_names(pok)
```

The variable names in the data set were all standardized. This function is useful because inconsistent formats for variables can lead to messy, less readable code and can potentially introduce bugs.

### Exercise 2

```{r}
type_1_plot<-ggplot(pok, aes(type_1)) + 
  geom_bar()
type_1_plot
```

```{r}
pok2 <- subset(pok, type_1 ==c('Bug','Fire','Grass',"Normal",'Water','Psychic'))
pok2$type_1 <- as.factor(pok2$type_1)
pok2$legendary <- as.factor(pok2$legendary)
```

There are 18 classes of the type_1 variable. There are very few Flying and Fairy types

### Exercise 3


```{r}
set.seed(619)

pok_split <- initial_split(pok2, prop = 0.80, strata = type_1)
pok_train <- training(pok_split)
pok_test <- testing(pok_split)
dim(pok_train)
dim(pok_test)
```

```{r}
pok_folds <- vfold_cv(pok_train, v = 5, strata = type_1)
pok_folds
```

Stratifying the folds would be useful becuause some predictor possibilities occur rarely, and ina randomly selected sample without stratification, some of these possibilities might get passed over entirely and fail to assist in training.

### Exercise 4


```{r}
pok_rec <- recipe(type_1 ~ legendary + generation + sp_atk + attack + speed + defense + hp + sp_def, data = pok_train) %>% 
  
  step_dummy(legendary, generation) %>% 
  step_normalize(all_numeric_predictors())
```

### Exercise 5

```{r}
net_mod <- 
  multinom_reg(penalty = tune(), mixture = tune()) %>% 
  set_mode('classification') %>% 
  set_engine('glmnet')

net_wkflw <- workflow() %>% 
  add_recipe(pok_rec) %>% 
  add_model(net_mod)
```


```{r}
net_grid <- grid_regular(penalty(range = c(-5, 5)), mixture(range = c(0,1)), levels = 10)
net_grid

```


I will be fitting 100 models to each of 5 folds, for a total of 500 model

### Exercise 6

```{r, eval = FALSE}
tune_res <-tune_grid(
  net_wkflw,
  resamples = pok_folds,
  grid = net_grid
)
```

```{r, eval = FALSE}
write_rds(tune_res, file = 'SavedModels/pok_tuned_res.rds')
```

```{r}
pok_tune_res <- read_rds(file = 'SavedModels/pok_tuned_res.rds')
```

```{r}
autoplot(pok_tune_res)
```

Medium to smaller values of penalty, and smaller values of mixture, produced better accuracy and ROC AUC.

### Exercise 7

Use `select_best()` to choose the model that has the optimal `roc_auc`. Then use `finalize_workflow()`, `fit()`, and `augment()` to fit the model to the training set and evaluate its performance on the testing set.

```{r}
best_mod <- select_best(pok_tune_res, metric = 'roc_auc')
pok_final <- finalize_workflow(net_wkflw, best_mod)
pok_final_fit <- fit(pok_final, data = pok_train)
```

```{r}
augment(pok_final_fit, new_data = pok_test) %>%
  conf_mat(truth = type_1, estimate = .pred_class) 
```



### Exercise 8

Calculate the overall ROC AUC on the testing set.

Then create plots of the different ROC curves, one per level of the outcome. Also make a heat map of the confusion matrix.

What do you notice? How did your model do? Which Pokemon types is the model best at predicting, and which is it worst at? Do you have any ideas why this might be?

```{r}
predicted_data <- predict(pok_final_fit, new_data = pok_test, type = 'prob')
```

```{r}
mod_roc_auc <- augment(pok_final_fit, new_data = pok_test) %>% 
 roc_a(truth = type_1, estimate = .pred_class)
```


```{r}
  roc_auc(type_1, .pred_Bug:.pred_Water) %>% 
  roc_curve(pok_train)

augment(pok_final_fit, new_data = pok_test) %>%
  conf_mat(truth = type_1, estimate = .pred_class) %>% 
  autoplot(type = 'heatmap')

```

