---
title: "DatasetMemo"
author: "Ben Hertzberg"
date: '2022-10-03'
output: pdf_document
---

##Overview of Dataset
The dataset I've chosen includes information on used car listings from dealerships around the Unites States. I downloaded the data from Kaggle: https://www.kaggle.com/datasets/3ea0a6a45dbd4713a8759988845f1a58038036d84515ded58f65a2ff2bd32e00?resource=download&select=ca-dealers-used.csv

The set includes over 7 million observations, but I will be truncating it to the first 100,000 observations so it is easier to handle. There are 21 variables associated with each observation. One is the price, which is the value I will attempt to predict. Some variables, such as dealership name or VIN, I will not use in the model. I decided to use 14 of the variables provided as predictors. There are both quantitative and qualitative predictor variables. The target variable is quantitative, which means this will be a regression model.

#check missing data

What variable(s) are you interested in predicting? What question(s) are you interested in answering?
Name your response/outcome variable(s) and briefly describe it/them.
Will these questions be best answered with a classification or regression approach?
Which predictors do you think will be especially useful?
Is the goal of your model descriptive, predictive, inferential, or a combination? Explain.

##Overview of Research Question
I am interested in predicting the listing price of a used car. Specifically, I want to know how what the price would be given information about the type and condition of the car. The response variable I will be using is called price in this data set, and provides the price that a certain used car was listed for at a dealership. My goal will be best achieved using a regression approach because the target variable is quantitative. I think the make, model, year, and mileage of each observation will be especially useful in predicting the price. The goal of my model is predictive, because I am not testing a theory or visually emphasizing a trend but rather attempting to predict a target variable given some data.

##Proposed Project Outline
I have my data set loaded already. A general timeline I'd like to follow is described below.
Week 1: Github Prep and Data choice
Week 2: Exploratory Data Analysis
Week 3: Data Splitting
Week 4 onward: Fit 4 models and narrate process
Week 10: Tune Best model

##Questions/Concerns


```{r}
usedcars <- us_dealers_used
```


```{r}
headuc = head(us_dealers_used, 10)
headuc
```

